Wrangling multivariate spatio-temporal data with the R package cubble

### 2022-04-30 arxiv/cubble-arxiv.tex
Updating abstract

\begin{abstract}
Multivariate spatio-temporal data refers to multiple measurements taken across space and time. For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time. However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance, temporal trends of multiple variables across locations. In order to facilitate the study of different sections or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components, for visual and numerical explorations while facilitating data wrangling for modelling. The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis. The \code{cubble} structure and the tools provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures. The tools are illustrated with examples of Australian climate data, merging climate and river level data sources, reproducing a climate reanalysis (ERA5), and creating interactive graphics.
\end{abstract}

@@

The SpatioTemporal CRAN task view (\protect\hyperlink{ref-ctvspatiotemporal}{Edzer Pebesma 2022}) gathers information about R packages designed for spatio-temporal data and it has a section on \emph{Representing data} that lists existing spatio-temporal data representations used in \proglang{R}. Among them, E. Pebesma (\protect\hyperlink{ref-spacetime}{2012}) summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats. The associated package \pkg{spacetime} (\protect\hyperlink{ref-spacetime}{E. Pebesma 2012}) implements four spatio-temporal layouts (full grid, sparse grid, irregular, and trajectory) to handle different space and time combinations. The \pkg{stars} (\protect\hyperlink{ref-stars}{E. Pebesma 2021}) package has been a recent implementation to use dense arrays to represent spatio-temporal cubes. It also interfaces with \pkg{sf} (\protect\hyperlink{ref-sf}{E. J. Pebesma 2018}), a package commonly used for wrangling spatial data, and the \pkg{tidyverse} (\protect\hyperlink{ref-tidyverse}{Wickham et al. 2019}) suite for general data wrangling and visualisation in \proglang{R}.

@@

This paper presents a new \proglang{R} package, \pkg{cubble} which addresses the two issues mentioned above. In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronised. Among the four spacetime layouts in E. Pebesma (\protect\hyperlink{ref-spacetime}{2012}), \pkg{cubble} can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work. The software is available from the Comprehensive R Archive Network (CRAN) at \url{https://CRAN.R-project.org/package=cubble}. \textbf{At this point what is cubble is adding with respect the other packages it is not clear to me}

@@

Spatio-temporal data can be conceptualised using a cubical data model with three axes, typically time, latitude and longitude. This abstraction can be useful for generalising operations and visualisation purposes: Lu, Appel, and Pebesma (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; Bach et al. (\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations. Notice that the term space-time cube in their article ``does not need to involve spatial data'', but refers to ``an abstract 2D substrate that is used to visualize data at a specific time''. Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation still applies to our spatio-temporal data manipulation and visualisation.

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.007}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cubble\_nested }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
\NormalTok{cubble\_nested}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cubble\_nested }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{rain\_day =} \FunctionTok{sum}\NormalTok{(ts}\SpecialCharTok{$}\NormalTok{prcp }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cubble\_long }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_nested }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{()}
\NormalTok{cubble\_long}
\end{Highlighting}
\end{Shaded}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# add station{-}based features in the nested form.}
\NormalTok{dt }\SpecialCharTok{|\textgreater{}} \FunctionTok{mutate}\NormalTok{(fabletools}\SpecialCharTok{::}\FunctionTok{features}\NormalTok{(ts, tmax, }\FunctionTok{list}\NormalTok{(}\AttributeTok{tmax\_mean =}\NormalTok{ mean, }\AttributeTok{tmax\_var =}\NormalTok{ var)))}
\end{Highlighting}
\end{Shaded}

@@

{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/illu-matching} \caption{An illustration of temporal matching in cubble. Three highest peaks are identified in each series and intervals are constructed on series \code{A}. Two peaks in series \code{a} fall into the intervals and hence the two series are considered to have two matches.}\label{fig:illu-matching}
\end{figure}

@@

{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}

@@

In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function. The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in Wickham et al. (\protect\hyperlink{ref-Wickham2012-yr}{2012})). The data can then be piped into \code{ggplot} to create the glyph map as:

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{covid }\SpecialCharTok{|\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{lga }\SpecialCharTok{|\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{historical\_tmax }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{filter}\NormalTok{(}\FunctionTok{between}\NormalTok{(stringr}\SpecialCharTok{::}\FunctionTok{str\_sub}\NormalTok{(id, }\DecValTok{7}\NormalTok{, }\DecValTok{8}\NormalTok{), }\DecValTok{46}\NormalTok{, }\DecValTok{90}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{filter}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1971}\SpecialCharTok{:}\DecValTok{1975}\NormalTok{, }\DecValTok{2016}\SpecialCharTok{:}\DecValTok{2020}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{group\_by}\NormalTok{(}\AttributeTok{month =}\NormalTok{ lubridate}\SpecialCharTok{::}\FunctionTok{month}\NormalTok{(date),}
         \AttributeTok{group =} \FunctionTok{as.factor}\NormalTok{(}\FunctionTok{ifelse}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\textgreater{}} \DecValTok{2015}\NormalTok{,}
                                  \StringTok{"2016 \textasciitilde{} 2020"}\NormalTok{, }\StringTok{"1971 \textasciitilde{} 1975"}\NormalTok{))) }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{tmax =} \FunctionTok{mean}\NormalTok{(tmax, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|\textgreater{}} \FunctionTok{filter}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(ts) }\SpecialCharTok{==} \DecValTok{24}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|\textgreater{}} \FunctionTok{unfold}\NormalTok{(latitude, longitude)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nsw\_vic }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{filter}\NormalTok{(NAME }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\StringTok{"Victoria"}\NormalTok{, }\StringTok{"New South Wales"}\NormalTok{))}

@@

\begin{figure}
\includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/sherryzhang/Documents/research/paper-cubble/figures/basic-manip} \caption{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia. On the top left corner is an insetted plot of station Cobar highlighted in the black box.}\label{fig:basic-manip}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{station\_nested }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{cluster =}\NormalTok{ ...)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ station\_nested }\SpecialCharTok{|\textgreater{}} \FunctionTok{switch\_key}\NormalTok{(cluster)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ cluster\_nested }\SpecialCharTok{|\textgreater{}} \FunctionTok{get\_centroid}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/basic-agg} \caption{Profile of aggregated precipitation from 639 weather stations in Australia. Subplot A shows the glyph map of the weekly averaged precipitation of each cluster. The group number is printed in the middle of the y minor axis and can be used as a reference line to read the magnitude. Subplot B shows the station membership of each cluster.}\label{fig:basic-agg}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/matching-map} \caption{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/matching} \caption{Matched weather stations and river gauges on the map (A) and across time (B). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The water level reflects the increase in precipitation.}\label{fig:matching}
\end{figure}

@@

Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in Hersbach et al. (\protect\hyperlink{ref-hersbach2020era5}{2020}). It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere. Readers interested in the analysis of this figure can refer to Hersbach et al. (\protect\hyperlink{ref-hersbach2020era5}{2020}), Simmons et al. (\protect\hyperlink{ref-simmons2020global}{2020}), and Simmons et al. (\protect\hyperlink{ref-simmons2005ecmwf}{2005}) for more details.

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/sherryzhang/Documents/research/paper-cubble/figures/netcdf} \caption{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020).}\label{fig:netcdf}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# data pre{-}processing}
\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}

\CommentTok{\# created SharedData instance for crosstalk}
\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{|\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}

\CommentTok{\# create the spatial and temporal view each with a ShareData instance}
\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}
\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# data pre{-}processing}
\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|\textgreater{}}\NormalTok{ ...}

>>>>

\begin{abstract}
Multivariate spatio-temporal data refers to multiple measurements taken across space and time. For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time. However for some studies, it is important to analyse different aspects of the spatio-temporal data simultaneouly, like for instance, temporal trends of multiple variables across locations. In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new data structure, \code{cubble}, with a suite of functions enabling easy slicing and dicing on the different components spatio-temporal components. The proposed \code{cubble} structure ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis.In addition, \code{cubble} facilitates visual and numerical explorations of the data while easing data wrangling and modelling. The \code{cubble} structure and the functions provided in the \code{cubble} R package equip users with the capability to handle hierarchical spatial and temporal structures. The \code{cubble} structure and the tools implemented in the package are illustrated with different examples of Australian climate data.
\end{abstract}

@@

The SpatioTemporal CRAN task view (\protect\hyperlink{ref-ctvspatiotemporal}{Edzer Pebesma 2022}) gathers information about R packages designed for spatio-temporal data and it has a section on \emph{Representing data} that lists existing spatio-temporal data representations used in \proglang{R}. Among them, \protect\hyperlink{ref-spacetime}{E. Pebesma} (\protect\hyperlink{ref-spacetime}{2012}) summarises spatio-temporal data into three forms: time-wide, space-wide, and long formats. The associated package \pkg{spacetime} (\protect\hyperlink{ref-spacetime}{E. Pebesma 2012}) implements four spatio-temporal layouts (full grid, sparse grid, irregular, and trajectory) to handle different space and time combinations. The \pkg{stars} (\protect\hyperlink{ref-stars}{E. Pebesma 2021}) package has been a recent implementation to use dense arrays to represent spatio-temporal cubes. It also interfaces with \pkg{sf} (\protect\hyperlink{ref-sf}{E. J. Pebesma 2018}), a package commonly used for wrangling spatial data, and the \pkg{tidyverse} (\protect\hyperlink{ref-tidyverse}{Wickham et al. 2019}) suite for general data wrangling and visualisation in \proglang{R}.

@@

This paper presents a new \proglang{R} package, \pkg{cubble} which addresses the two issues mentioned above. In the package, a new data structure, also called \code{cubble}, is proposed to organise spatial and temporal variables as two forms of a single data object so that they can be wrangled separately or combined while being kept synchronised. Among the four spacetime layouts in \protect\hyperlink{ref-spacetime}{E. Pebesma} (\protect\hyperlink{ref-spacetime}{2012}), \pkg{cubble} can be applied to full grid, sparse grid, or irregular, but not trajectory, which is outside the scope of this work. The software is available from the Comprehensive R Archive Network (CRAN) at \url{https://CRAN.R-project.org/package=cubble}. \textbf{At this point what is cubble is adding with respect the other packages it is not clear to me}

@@

Spatio-temporal data can be conceptualised using a cubical data model with three axes, typically time, latitude and longitude. This abstraction can be useful for generalising operations and visualisation purposes: \protect\hyperlink{ref-lu_multidimensional_2018}{Lu, Appel, and Pebesma} (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; \protect\hyperlink{ref-bach_review_2014}{Bach et al.} (\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations. Notice that the term space-time cube in their article ``does not need to involve spatial data,'' but refers to ``an abstract 2D substrate that is used to visualize data at a specific time.'' Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation still applies to our spatio-temporal data manipulation and visualisation.

@@

{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.007}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cubble\_nested }\OtherTok{\textless{}{-}}\NormalTok{ climate\_flat }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
  \FunctionTok{as\_cubble}\NormalTok{(}\AttributeTok{key =}\NormalTok{ id, }\AttributeTok{index =}\NormalTok{ date, }\AttributeTok{coords =} \FunctionTok{c}\NormalTok{(long, lat))}
\NormalTok{cubble\_nested}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cubble\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{rain\_day =} \FunctionTok{sum}\NormalTok{(ts}\SpecialCharTok{$}\NormalTok{prcp }\SpecialCharTok{!=} \DecValTok{0}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cubble\_long }\OtherTok{\textless{}{-}}\NormalTok{ cubble\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{()}
\NormalTok{cubble\_long}
\end{Highlighting}
\end{Shaded}

@@

{\centering \includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.008}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# add station{-}based features in the nested form.}
\NormalTok{dt }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{mutate}\NormalTok{(fabletools}\SpecialCharTok{::}\FunctionTok{features}\NormalTok{(ts, tmax, }\FunctionTok{list}\NormalTok{(}\AttributeTok{tmax\_mean =}\NormalTok{ mean, }\AttributeTok{tmax\_var =}\NormalTok{ var)))}
\end{Highlighting}
\end{Shaded}

@@

{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.003}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/illu-matching} \caption{An illustration of temporal matching in cubble. Three highest peaks are identified in each series and intervals are constructed on series \code{A}. Two peaks in series \code{a} fall into the intervals and hence the two series are considered to have two matches.}\label{fig:illu-matching}
\end{figure}

@@

{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/diagram-keynotes/diagram-keynotes.004}

@@

In \proglang{R}, \pkg{GGally} implements glyph maps through the \code{glyphs()} function. The function constructs a data frame with calculated position (\code{gx}, \code{gy}, \code{gid}) of each point on the time series using linear algebra (Equation 1 and 2 in \protect\hyperlink{ref-Wickham2012-yr}{Wickham et al.} (\protect\hyperlink{ref-Wickham2012-yr}{2012})). The data can then be piped into \code{ggplot} to create the glyph map as:

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{covid }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{lga }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{head}\NormalTok{(}\DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{historical\_tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
  \FunctionTok{filter}\NormalTok{(}\FunctionTok{between}\NormalTok{(stringr}\SpecialCharTok{::}\FunctionTok{str\_sub}\NormalTok{(id, }\DecValTok{7}\NormalTok{, }\DecValTok{8}\NormalTok{), }\DecValTok{46}\NormalTok{, }\DecValTok{90}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
  \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
  \FunctionTok{filter}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\DecValTok{1971}\SpecialCharTok{:}\DecValTok{1975}\NormalTok{, }\DecValTok{2016}\SpecialCharTok{:}\DecValTok{2020}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
  \FunctionTok{group\_by}\NormalTok{(}\AttributeTok{month =}\NormalTok{ lubridate}\SpecialCharTok{::}\FunctionTok{month}\NormalTok{(date),}
         \AttributeTok{group =} \FunctionTok{as.factor}\NormalTok{(}\FunctionTok{ifelse}\NormalTok{(lubridate}\SpecialCharTok{::}\FunctionTok{year}\NormalTok{(date) }\SpecialCharTok{\textgreater{}} \DecValTok{2015}\NormalTok{,}
                                  \StringTok{"2016 \textasciitilde{} 2020"}\NormalTok{, }\StringTok{"1971 \textasciitilde{} 1975"}\NormalTok{))) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{tmax =} \FunctionTok{mean}\NormalTok{(tmax, }\AttributeTok{na.rm =} \ConstantTok{TRUE}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_spatial}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{filter}\NormalTok{(}\FunctionTok{nrow}\NormalTok{(ts) }\SpecialCharTok{==} \DecValTok{24}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tmax }\OtherTok{\textless{}{-}}\NormalTok{ tmax }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{face\_temporal}\NormalTok{() }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{unfold}\NormalTok{(latitude, longitude)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nsw\_vic }\OtherTok{\textless{}{-}}\NormalTok{ ozmaps}\SpecialCharTok{::}\NormalTok{abs\_ste }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
  \FunctionTok{filter}\NormalTok{(NAME }\SpecialCharTok{\%in\%} \FunctionTok{c}\NormalTok{(}\StringTok{"Victoria"}\NormalTok{, }\StringTok{"New South Wales"}\NormalTok{))}

@@

\begin{figure}
\includegraphics[width=1\linewidth,height=0.7\textheight]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/basic-manip} \caption{A glyph map of the average maximum temperature by month of Victoria and New South Wales weather stations in Australia. On the top left corner is an insetted plot of station Cobar highlighted in the black box.}\label{fig:basic-manip}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{station\_nested }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{cluster =}\NormalTok{ ...)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ station\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{switch\_key}\NormalTok{(cluster)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cluster\_nested }\OtherTok{\textless{}{-}}\NormalTok{ cluster\_nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}} \FunctionTok{get\_centroid}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/basic-agg} \caption{Profile of aggregated precipitation from 639 weather stations in Australia. Subplot A shows the glyph map of the weekly averaged precipitation of each cluster. The group number is printed in the middle of the y minor axis and can be used as a reference line to read the magnitude. Subplot B shows the station membership of each cluster.}\label{fig:basic-agg}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/matching-map} \caption{Location of weather stations and river gauges in Victoria, Australia.}\label{fig:matching-map}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/matching} \caption{Matched weather stations and river gauges on the map (A) and across time (B). Precipitation and water level have been standardised between 0 and 1 to be displayed on the same scale. The water level reflects the increase in precipitation.}\label{fig:matching}
\end{figure}

@@

Figure \ref{fig:netcdf} reproduces the ERA5 data row of Figure 19 in \protect\hyperlink{ref-hersbach2020era5}{Hersbach et al.} (\protect\hyperlink{ref-hersbach2020era5}{2020}). It shows the southern polar vortex splits into two on 2002-09-26 and further splits into four on 2002-10-04 in the stratosphere. Readers interested in the analysis of this figure can refer to \protect\hyperlink{ref-hersbach2020era5}{Hersbach et al.} (\protect\hyperlink{ref-hersbach2020era5}{2020}), \protect\hyperlink{ref-simmons2020global}{Simmons et al.} (\protect\hyperlink{ref-simmons2020global}{2020}), and \protect\hyperlink{ref-simmons2005ecmwf}{Simmons et al.} (\protect\hyperlink{ref-simmons2005ecmwf}{2005}) for more details.

\begin{figure}
\includegraphics[width=1\linewidth]{/Users/pmen0008/Google Drive (2)/ Private/Projects/paper-cubble/figures/netcdf} \caption{A reproduction of the second row (ERA5 data) of Figure 19 in Hersbach et al (2020).}\label{fig:netcdf}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# data pre{-}processing}
\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}

\CommentTok{\# created SharedData instance for crosstalk}
\NormalTok{nested }\OtherTok{\textless{}{-}}\NormalTok{ clean }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}
\NormalTok{long }\OtherTok{\textless{}{-}} \FunctionTok{face\_temporal}\NormalTok{(clean) }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ SharedData}\SpecialCharTok{$}\FunctionTok{new}\NormalTok{(}\SpecialCharTok{\textasciitilde{}}\NormalTok{id, }\AttributeTok{group =} \StringTok{"cubble"}\NormalTok{)}

\CommentTok{\# create the spatial and temporal view each with a ShareData instance}
\NormalTok{p1 }\OtherTok{\textless{}{-}}\NormalTok{ nested }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
\NormalTok{p2 }\OtherTok{\textless{}{-}}\NormalTok{ long }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# data pre{-}processing}
\NormalTok{clean }\OtherTok{\textless{}{-}}\NormalTok{ weatherdata}\SpecialCharTok{::}\NormalTok{climate\_full }\SpecialCharTok{|}\ErrorTok{\textgreater{}}\NormalTok{ ...}
